---
title: "DEC Data Assembly Compiled"
output: html_notebook
---
Authors: Grace Kuiper, Nelsha Athauda
Email: grace.kuiper@colostate.edu, nrathauda@alaska.edu


```{r Here, include=FALSE, message=F}
here::i_am("scripts/DEC Data Assembly Compiled.Rmd")
knitr::opts_knit$set(root.dir = here::here())
```

```{r Libraries}
library(tidyverse)
library(lubridate)
library(dplyr)
library(here)
library(stringr)
```

# File Upload
```{r Read Data}
raw_V2_data <- read.csv(here("raw_data", "AMP435_2094888-0.txt"))
```

# Remove bottom line descriptor, rename State Code, and remove space at the end of County Names
```{r}
DEC_data <- raw_V2_data %>%
  head(-1) %>%
  rename(STATE.CODE = X..STATE.CODE) %>%
  mutate(COUNTY.NAME = str_trim(COUNTY.NAME, side = "right")) 

table(DEC_data$COUNTY.NAME)
table(DEC_data$SITE.ID)
```
# Format date variable
```{r}
DEC_data$COLLECTION.DATE <- as.character(DEC_data$COLLECTION.DATE)
DEC_data$COLLECTION.DATE <- as.Date(DEC_data$COLLECTION.DATE, origin = "20000101", tryFormats = c("%Y%m%d"))
DEC_data$Date <- DEC_data$COLLECTION.DATE
DEC_data = DEC_data[,!(names(DEC_data) %in% c("COLLECTION.DATE"))]
```

# Eliminate readings with fewer than 75% of measurements (18 hours) for summarized hourly data
```{r}
avg_DEC_data <- DEC_data %>%
  filter(`PCT.DAILY.OBS`>=75) %>%
  filter(ifelse(`DURATION.DESC`=="1 HOUR",`NUM.DAILY.OBS`>=18,`NUM.DAILY.OBS`>0)) %>%
  filter(`ARITHMETIC.MEAN`>0)
```

# Create Monitor ID Tags
```{r}
avg_DEC_data$STATE.CODE<- as.numeric(avg_DEC_data$STATE.CODE)
avg_DEC_data$COUNTY.CODE<- as.numeric(avg_DEC_data$COUNTY.CODE)
avg_DEC_data$SITE.ID<- as.numeric(avg_DEC_data$SITE.ID)

avg_DEC_data <- avg_DEC_data %>%
  mutate(MONITOR.ID = paste(STATE.CODE, COUNTY.CODE, SITE.ID, sep = "-"))

table(avg_DEC_data$MONITOR.ID)
```

# Add in long/lat for monitor sites from EPA Website (WGS84)
```{r}
avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-110-26"] <- "-134.510427" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-110-26"] <- "58.358875"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-110-27"] <- "-134.516499" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-110-27"] <- "58.365508"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-110-4"] <- "-134.567237"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-110-4"] <- "58.388497"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-122-8"] <- "-151.069300"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-122-8"] <- "60.489167"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-130-8"] <- "-131.650019"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-130-8"] <- "55.351043"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-170-12"] <- "-149.106220" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-170-12"] <- "61.598898"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-170-13"] <- "-149.455255" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-170-13"] <- "61.598796"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-170-4"] <- "-150.252229" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-170-4"] <- "61.532508"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-170-8"] <- "-149.031655" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-170-8"] <- "61.534163"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-20-1004"] <- "-149.569707" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-20-1004"] <- "61.326700"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-20-18"] <- "-149.824602" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-20-18"] <- "61.205861"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-20-44"] <- "-149.817389" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-20-44"] <- "61.181083"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-20-52"] <- "-149.903111" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-20-52"] <- "61.215027"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-232-1"] <- "-135.321250" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-232-1"] <- "59.453150"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-232-2"] <- "-135.310530" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-232-2"] <- "59.460080"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-68-3"] <- "-148.967600" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-68-3"] <- "63.723200"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-10"] <- "-147.722461"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-10"] <- "64.840672"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-29"] <- "-147.692461"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-29"] <- "64.846506"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-33"] <- "-147.347208" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-33"] <- "64.752336"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-34"] <- "-147.727413" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-34"] <- "64.845690"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-35"] <- "-147.310279"
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-35"] <- "64.762641"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-39"] <- "-147.372278" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-39"] <- "64.759289"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-40"] <- "-147.693270" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-40"] <- "64.845930"

avg_DEC_data$Longitude[avg_DEC_data$MONITOR.ID == "2-90-4003"] <- "-147.872000" 
avg_DEC_data$Latitude[avg_DEC_data$MONITOR.ID == "2-90-4003"] <- "64.839130"


avg_DEC_data$Longitude <- avg_DEC_data$Longitude
avg_DEC_data$Latitude <- avg_DEC_data$Latitude
```

# Create Priority Variable
```{r}
avg_DEC_data$Priority <- NA

avg_DEC_data$Priority[avg_DEC_data$DURATION.DESC == "1 HOUR"] <- 1
avg_DEC_data$Priority[avg_DEC_data$DURATION.DESC == "24 HOUR"] <- 2
avg_DEC_data$Priority[avg_DEC_data$DURATION.DESC == "24-HR BLK AVG"] <- 3

avg_DEC_data$Priority <- avg_DEC_data$Priority
```

# Select one measurement per site per day by Priority
```{r}
if (anyDuplicated(avg_DEC_data[, c("COUNTY.NAME", "SITE.ID", "Date", "MONITOR.ID")])) {
  avg_DEC_data <- avg_DEC_data %>%
    group_by(COUNTY.NAME, SITE.ID, Date, Latitude) %>%
    mutate(has_duplicate = n() > 1) %>%
    arrange(has_duplicate, Priority) %>%
    filter(!has_duplicate | row_number() == 1) %>%
    ungroup() %>%
    select(-has_duplicate)
}

dim(avg_DEC_data)
```

```{r}
write.csv(avg_DEC_data, here("clean_data", "Cleaned_DEC_Data_05302023.csv"))
```



